
```{r setup, include = FALSE}
library(rmarkdown)
library(purrr)
library(googledrive)

knitr::opts_chunk$set(
  include = FALSE
)

```


Hi, Maggie,

Run this script if you need to rebuild the rubrics. I've put them in both
`word_document` and `pdf_document` format. Just hit the "source" button.
Hopefully you wont need this often; it's mostly to have rubrics with embedded
plots that can be changed easily.   Make sure your rubrics are named like
_RM1_Rubric.Rmd, _RM2_Rubric.Rmd, etc. The leading underscore tells R studio
to ignore the file when building the website

``` {r}
rubric_list <- list.files(pattern = "^_.+_Rubric\\.Rmd")

map(
  rubric_list,
  function(input) {
    rmarkdown::render(
      input = input, output_dir = "rubrics/docx",
      output_format = "word_document"
    )
  }
)

map(
  rubric_list,
  function(input) {
    rmarkdown::render(
      input = input, output_dir = "rubrics/pdf",
      output_format = "pdf_document"
    )
  }
)
```

This function uploads the .docx files to Google Drive and this chunk needs to be run manually
``` {r, eval = FALSE}
rubric_docx <- list.files(
  path = "rubrics/docx",
  pattern = "^_.+\\.docx",
  full.names = TRUE)

map(rubric_docx, function(x) {
  base <- basename(x)
  num <- str_extract(x, "\\d+")
  name <- paste0("R Module ", num)
  dir <- paste0("Updated R Modules/", name, "/")
  drive_upload(x, path = dir, name = paste0(name, " Rubric.docx"), overwrite = T)
})
```