```{r setup, include=F}
knitr::opts_chunk$set(out.width = "50%",
                      message = FALSE)
```

# R Module 1 Rubric

This R module is pretty simple; there aren't that many difficult questions. Students should mostly just be able to load and run R Studio.

```{r}
# Load data
library(MASS)

head(painters)
```

**Use R to produce the following simple data summaries and visualizations commonly used for qualitative data:**

## Question 1:

*A histogram of the `Composition` variable from the `painters` dataset, using the `hist()` function.*

```{r q1}
hist(painters$Composition)
```


## Question 2:

*A bar chart of the `Expression` variable, using the `barplot()` function.*

    -   *Tip: use \`?barplot() to see which argument you need to change to get a horizontal bar plot* Also, wrap the painters$Expression with the table() function, then use the barplot() function

```{r q2}
barplot(table(painters$Expression))
```


## Question 3: 

*A pie chart of the `Schools` variable, using the `pie()` function.*

    -   *Hint: you need to wrap the `pie()` function around the `table()` function, because you need **counts** of each school, rather than just their label*

```{r q3}
pie(table(painters$School))
```


## Question 4:

*A histogram for the eruptions. R will automatically decide how to group your observations. Remember to plot __only__ the `eruptions` variable.*

```{r q4}
hist(faithful$eruptions)
```


## Question 5:

*Another histogram of eruptions, but specify 2 data classes instead of the default values.*

```{r q5}
hist(faithful$eruptions, nclass = 2)
```

## Question 6:

*A boxplot of eruptions that is horizontal, rather than vertical.*

```{r q6}
boxplot(faithful$eruptions,
        horizontal = T)
```


## Question 7: 

*Repeat steps 4-6 (histogram, histogram with nbreaks = 2, and horizontal boxplot) on a dataset of your own choosing and interpret the data. You can also use functions such as `mean`, `median`, `quantile`, etc. to help interpret your data. Give each of your new plots x- and y-axis labels and a title*

```{r q7}
mydata <- mtcars$mpg

# Students don't have to use the par(mfrow); I've just used it to tidy the plots
# a bit.

par(mfrow = c(2, 2))

hist(mydata,
     xlab = "MPG",
     ylab = "Frequency",
     main = "Histogram of MPG")

hist(
  mydata,
  nclass = 2,
  xlab = "MPG",
  ylab = "Frequency",
  main = "Histogram of MPG"
)

boxplot(mydata,
        horizontal = TRUE,
        xlab = "MPG",
        main = "Boxplot of MPG")

par(mfrow = c(1, 1))
```
