```{r setup, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  out.width = "50%"
)
```

#R Module 2: Working with R Scripts and Projects

This R Module is intended to introduce summary statistics, measures of central tendency, and measures of dispersion. In terms of R concepts, it will teach students about how to work with files in the working directory, create and use scripts (rather than type all the code in the console!) and organize their R Modules into an R Project (.rproj).

```{r}
# Load in the data
library(datasets)
tree_data <- trees
```



1.  Create your own histogram of the `tree_data` dataset with `ggplot2`, showing:
    - Distribution of the `Height` variable,
    - 20 bins,
    - A color of your choosing,
    - Labels (title and x-axis) that reflect the change in variable


```{r q1}
library(ggplot2)
ggplot(
  data = tree_data,
  aes(x = Height)
) +
  geom_histogram(bins = 20, fill = "brown3") +
  labs(
    x = "Height",
    y = "Frequency",
    title = "Histogram of Tree Height"
  )
```

Question 2:

Visualize and Explore the `mtcars` dataset:
1.  Create a new script called `mtcars.R`
2. Assign `mtcars` to an object (e.g., “cars”) using the <- operator
3. Choose 3 variables from `mtcars` and for each:
    - Record the Mean, Median, and Mode
    - Create a histogram in `ggplot2`, with appropriate labels, title, etc.
    
```{r q2}
# The code for this one should include their mtcars.R script as well

cars <- mtcars

# They should have three of the following, for each of the variables they choose:
mean(mtcars$mpg)

median(mtcars$mpg)


# Any way they want to find the mode is fine. Oddly, base R doesn't have a function to find the mode.
find_mode <- function(x) {
  unique_data <- unique(x)
  unique_data[which.max(tabulate(match(x, unique_data)))]
}

find_mode(mtcars$mpg)

```

```{r q2-3-2}
ggplot(mtcars, aes(x = mpg)) + 
  geom_histogram() + 
  labs(
    x = "Miles per Gallon",
    y = "Count",
    title = "Histogram of mpg"
  )
```


d. Choose two variables from `mtcars` and create a scatter plot in `ggplot2` showing their relation:
    - Include title, labels, etc.
    - *Hint: Use the geom_point() function*
    - Use Google, Stack Exchange, and the built-in Help for examples on how to use these plots!
    

```{r q2-4}

ggplot(data = mtcars, aes(x = mpg, y = hp)) + 
  geom_point() + 
  labs(x = "Miles per Gallon",
       y = "Horsepower",
       title = "mpg. vs. hp.")

```




Include a screenshot of your `mtcars.R` script
Submit all questions to Google Classroom; include Name, Question Number, and all figures
