```{r setup, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  out.width = "50%"
)
```

# R Module 5 Rubric

## Questions

1. Adapt the code above to calculate the area under the curve (probability) for the following intervals of `x`:
    - (-2, 2)
    - (-3, 3)
    - (-4, 4)

```{r}
pnorm(2) - pnorm(-2)
pnorm(3) - pnorm(-3)
pnorm(4) - pnorm(-4)
```




2. Use `qnorm()` to calculate the values of `x` for each of the following circumstances. Provide both the calculated values of `x` and your R code.
  a. 20% of the area of the curve is to the left of an unknown value of x
  b. 67% of the area of the curve is to the right of an unknown value of x

```{r}
qnorm(0.2)
1 - qnorm(0.67)
```

\newpage

3. Use the salary data from above to create 90%, 95%, and 99% confidence intervals. Provide each interval, along with your code. What happens to the interval range as you increase the confidence percentage?



```{r q3}

x <- c(44617, 7066, 17594, 2726, 1178, 18898, 5033, 37151, 4514, 4000)

confidence <- function(x, sdev = 15000, ci) {
  n <- length(x)
  xbar <- mean(x)
  alpha <- (1 - ci) / 2
  z.alpha.2 <- qnorm(p = 1 - alpha,
                     mean = 0,
                     sd = 1)
  lbound <-
    xbar - (z.alpha.2 * (sdev / sqrt(n)))
  ubound <-
    xbar + (z.alpha.2 * (sdev / sqrt(n)))
  
  lbound <- round(lbound, digits = 2)
  ubound <- round(ubound, digits = 2)
  
  return(c(lbound, ubound))
}

for (i in c(0.9, 0.95, 0.99)) {
  print(confidence(x, ci = i))
}

```

*As we increase the confidence level, the confidence interval widens, because we need a larger range to be confident that the population mean lies within our range*
