## Setup

```{r, warning=FALSE}
library(sf)
library(tidyverse)
library(spdep)
library(tmap)
library(broom)
library(spatialreg)

election <- read_sf("export/2004_Election_Counties.shp") %>%
  st_transform(crs = st_crs("EPSG:2163"))

centroids <- st_centroid(election)$geometry

election_neighbors <- election %>%
  poly2nb()

election_weighted <- election_neighbors %>%
  nb2listw(
    style = "W",
    zero.policy = TRUE
  )
```


## Question 1:

Use `lm()` to estimate `Bush_pct ~ pcincome` using ordinary least squares (OLS). Plot the **residuals** on a map (look back to R Module 7) and **test the residuals for spatial autocorrelation**. Provide both an example of your map and the results from your spatial autocorrelation test in your R Markdown report. 

```{r q1-1}
lm <- lm(Bush_pct ~ pcincome, data = election)
summary(lm)
```

```{r q1-2}
election$residuals <- lm$residuals

tm_shape(election) +
  tm_polygons(
    col = "residuals",
    border.col = NULL,
    style = "cont"
  )
```

```{r q1-3}
moran_residuals <-
  moran.test(
    x = election$residuals,
    listw = election_weighted,
    zero.policy = TRUE,
    alternative = "two.sided"
  )

moran_residuals
```


## Question 2:

Modeling `Kerry_pct ~ pcincome`, perform the following steps:
  - Evaluate the relationship between the variables using `lm()`
  - Plot the resulting residuals, then test for autocorrelation with `moran.test()`. Provide your map!



```{r q2-1}

kerry_lm <- lm(Kerry_pct ~ pcincome, data = election)

election$kerry_lm_res <- kerry_lm$residuals

tm_shape(election) +
  tm_polygons(
    col = "kerry_lm_res",
    border.col = NULL,
    palette = "RdBu",
    style = "cont",
    midpoint = 0,
    title = "Residuals"
  )
```


```{r q2-2}
kerry_moran <- moran.test(
  election$kerry_lm_res,
  listw = election_weighted,
  zero.policy = TRUE,
  alternative = "two.sided"
)
kerry_moran
```

  - Estimate the same relationship, using the spatial lag model (`spatialreg::lagsarlm()`)
  
```{r q2-3, eval = FALSE}
library(spatialreg)
kerry_lag <- lagsarlm(
  Kerry_pct ~ pcincome,
  data = election,
  listw = election_weighted,
  zero.policy = TRUE
)

tidy(kerry_lag)
```
  
```{r, include = FALSE}

kerry_lag <- read_rds("export/kerry_lag.Rds")

tidy(kerry_lag)
```
  
  
  - Perform a Moran's I test on the residuals from the lag model and report your results. Plot the residuals from the spatial lag model, and include in your report.
  
```{r q2-4}
election$kerry_lag_res <- kerry_lag$residuals

tm_shape(election) +
  tm_polygons(
    col = "kerry_lag_res",
    border.col = NULL,
    palette = "RdBu",
    style = "cont",
    midpoint = 0,
    title = "Residuals"
  )
```


```{r q2-4b}
kerry_lag_moran <- moran.test(
  election$kerry_lag_res,
  election_weighted, ,
  zero.policy = TRUE,
  alternative = "two.sided"
)
kerry_lag_moran
```
  
  - Is there evidence of remaining spatial autocorrelation in the residuals?
  
  *No, as the value of I from the moran.test is close to 0, there is little remaining spatial autocorrelation.*
