---
title: "R Markdown Reports and Survey Data"
output:
  html_document:
      toc: true
      toc_float: true
      toc_collapsed: true
      toc_depth: 3
      number_sections: false
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(
  echo = FALSE
)

library(tidyverse)


airport_survey <- read_csv("data/airport_survey.csv", 
    col_types = cols(`Date recorded` = col_date(format = "%m/%d/%Y")))


```



# Generating Reports using R Markdown

For this R Module, we'll be taking on the role of ...

## What is Markdown?

Markdown is a popular "markup language" designed to format plain-text data, such as comments on many websites and documentation for code. In text editors like Microsoft Word or Google Docs, called "What you see is what you get (*WYSIWYG*)" editors, there are buttons for making text italic, bold, underlined, etc., but it's often difficult to transfer that information from one editor to another; if you copy italic text from a Word document into a program like Notepad, you lose all your formatting. 

Markdown, however is designed to be a "What you type is what you get (*WYTIWYG*)" language. If you want to *italic* text, you wrap your text with single \*asterisks\* or \_underscores\_. If you want **bold** text, you \*\*use\*\* \_\_two\_\_, and you can ***combine*** them with \*\*\*three\*\*\*.

There are a ton of other formatting options available, such as lists, links, etc. This module isn't intended to teach Markdown, but rather a variant called R Markdown. For more information regarding Markdown, you can use sites like [Handbook's Markdown Guide](https://about.gitlab.com/handbook/markdown-guide/) or resources such as Stack Overflow.

## R Markdown

R Markdown is a variant of Markdown that is designed to work with R code. In addition to all of the standard Markdown formatting rules, R Markdown allows you to embed and run R code directly in your document, in sections called *code chunks*. This lets you not only run your code, but generate a document with text as well; perfect for writing up labs! Once you get the hang of R Markdown documents, you won't want to write up reports in Word anymore. Rather than endlessly take screenshots and paste into a document, you can embed your code in an R Markdown document, and it'll insert the plots and graphs that you generate *directly*, like this:

<pre><code>```{r code-example}

x <- seq(-1, 1, length = 20)

y <- dnorm(x)

plot(x, y)

```</code></pre>

Which gives you:


```{r code-example, echo = FALSE, out.width = "50%"}
x <- seq(-1, 1, length = 20)
y <- dnorm(x)

plot(x, y)

```

Or, if you want to print a table, you can use:

<pre><code>```{r table}

head(mtcars)

```</code></pre>

Which gives:

```{r table, echo = FALSE}

head(mtcars)

```


# Getting Started

Before we do anything with R Markdown, we need to do a bit of setup to make sure that R Studio is good-to-go in a few ways. 

First, if you haven't already, install *R Tools*. This software is used when installing packages, and makes things a lot easier. For Windows, you can navigate to:
  https://cran.r-project.org/bin/windows/Rtools/rtools40.html

Download the 64-bit version, and install the software. Note that this requires admin priviledges. This step isn't strictly necessary; R tools is used to build packages that are "in development", but it's useful to have if you can.

After that, we'll need to install a few R packages:

  - `rmarkdown`, `knitr`, `markdown`, and `mime`: These allow R Studio to work with Markdown and R Markdown
  - `tinytex`: This package is a lightweight $\LaTeX$ distribution and is necessary to export as a .PDF
    - After installing this package, run `tinytex::install_tinytex()`. This step might take some time. 
  

To get things running, you can just run the following code:

```{r packages, eval = FALSE}
pkgs <- c("rmarkdown", "knitr", "markdown", "mime", "tinytex")

install.packages(pkgs)

tinytex::install_tinytex()

```

Then, restart R Studio.


Let's get started by creating a new R Project, as in previous R Modules. Once you have that set up, we want to create a new R Markdown document. Navigate to `File > New File > R Markdown`. Give it a title, "R Module 4 Writeup", enter your name, and select `PDF` as the output format. 



### Running Code



# Qualitative Data

### Factors and Levels in R

***

## Recoding using `dplyr`


***

## Visualizing Qualitative Data with `ggplot2`

***
***


# Survey Development, Sampling, and Statistical Tests

## Sampling Techniques

***


## Statistisal Tests for Surveys

***
***


# Assignment